---
title: "Lab 2"
author: "Daniel Anderson"
date: "4/10/2019"
output:
  html_document: 
    toc: true
    toc_float: true
    theme: "journal"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This lab will be similar to Lab 1 but will focus specifically on looping with 
{purrr}.


### Part A: Multiple Models

Run the code below to load the following dataset.

```{r load_data}
file <- "https://github.com/datalorax/esvis/raw/master/data/benchmarks.rda"
load(url(file))
head(benchmarks)
```

These data are simulated, but represent individual student scores across seasonal benchmark screenings (administered in the fall, winter, and spring). The variables are as follows:

* `sid`: student identifier
* `cohort`: student cohort
* `sped`: special education status of the student for the given time point
* `ethnicity`: coded race/ethnicity of the student
* `frl`: free or reduced price lunch eligibility for the student for the given time point
* `ell`: English language learner status for the student for the given time point
* `season`: season the assessment was administered
* `reading`: student's reading score at the given time point
* `math`: student's mathematics score at the given time point

Recode season to wave with the following code

```{r recode, message = FALSE, warning = FALSE}
library(tidyverse)

benchmarks <- benchmarks %>%
	as_tibble() %>%
	mutate(wave = case_when(season == "Fall" ~ 0,
	                        season == "Winter" ~ 1,
	                        TRUE ~ 2))
```

1. Fit a model of the form `lm(math ~ wave)` for each student. 

2. Plot the distribution of slopes. Annotate the plot with a vertical line for the mean.

```{r answer1, include = FALSE}
slopes <- split(benchmarks, benchmarks$sid) %>%
	map(~lm(math ~ wave, data = .x)) %>%
	map(coef) %>%
	map_dbl(2)

to_plot <- tibble(sid = names(slopes),
                  slope = slopes)

theme_set(theme_minimal())
ggplot(to_plot, aes(slopes)) +
	geom_histogram(fill = "cornflowerblue",
	               color = "white") +
	geom_vline(xintercept = mean(slopes, na.rm = TRUE),
	           color = "magenta",
	           lwd = 2)
```

### Part B: Star Wars

Install and load the {repurrrsive} package

```{r starwars-data, eval = FALSE}
install.packages("repurrrsive")
library(repurrrsive)
```

You now should have access to all the data you'll need for the following.

1. Use the `sw_films` list to identify how many characters were represented in each film.

2. Use the `sw_species` list to (a) identify species that have known hair colors and (b) identify what those hair colors are.

```{r starwars-answers, include = FALSE}
library(repurrrsive)
titles <- map_chr(sw_films, "title")
n_char <- map(sw_films, "characters") %>%
	map_dbl(length) %>%
	setNames(titles)

tibble(species = map_chr(sw_species, "name"),
       hair = map_chr(sw_species, "hair_colors")) %>%
	filter(hair != "none" & hair != "n/a" & hair != "unknown")
```


### Part C: Some basic API calls

Many of the list elements have urls for API calls. For example

```{r people1}
sw_films[[1]]$characters[1]
```

The above provides the url necessary to make an API call for all the information available on the first character in the first film. You can get this information into R with the following code

```{r char1, eval = FALSE}
library(httr)
char_1 <- GET(sw_films[[1]]$characters[1]) %>%
	content("parsed") 
```

You'll note it looks very similar to the lists we've been working with. If you're interested, you may want to actually visit the link in your browser. Essentially, the call requests that the API provide the output available for the first character. This is returned in JSON format, and the `content(x, "parsed")` part transforms it into a nested list.

1. Create a data frame that has two columns listing the names of every character in the first film (i.e., all the characters in `sw_films[[1]]$characters`), and the total number of films each character appeared in (note, this may take a minute or two).


```{r part-c-1, include = FALSE, eval = FALSE}
library(httr)
chars <- map(sw_films[[1]]$characters, GET)
char_info <- map(chars, content, "parsed")

char_tbl <- tibble(character = map_chr(char_info, "name"),
                   n_films = map(char_info, "films") %>% map_int(length))
```

#### Bonus: Up to 1 point extra credit

2. Add an additional column to your data frame that includes the names of all the films each character was in. 

```{r part-c-2, include = FALSE, eval = FALSE}
films <- map(char_info, "films")
film_info <- map(films, ~map(.x, GET))
char_films <- map(film_info, ~map(.x, content, "parsed")) %>%
	map(~map_chr(.x, "title"))

char_tbl <- char_tbl %>%
	mutate(film_titles = map_chr(char_films, paste0, collapse = ", "))
```

