---
title: "Lab 3"
author: "Daniel Anderson"
date: "Assigned: 5/8/19; Due 5/15/19"
output:
  html_document: 
    toc: True
    toc_float: true
    theme: "journal"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This lab focuses on writing basic functions.

### Part A: Foundations
1. Briefly name and describe the three fundamental components of a function.

```{r ans1a, include = FALSE}
# 1. Formals - The arguments supplied to the function
# 2. Body - What the function actually does, including the use of the formals
# 3. Environment - Where the function "lives", which is important for scoping.
```

2. Describe three different ways functions can be stored/applied and an example
of when you might want to use each version.

```{r ans2b, include = FALSE}
# anonymous functions are often used in loops, usually functionals like
# `purrr::map` or lapply.

# Function are usually bound to a name and applied like we typically see
# functions, with the name preceding parentheses (e.g., `lm()`)

# Functions can also be bound to names within lists, which can be useful if you
# want to loop a set of functions through something, rather than looping a
# single function through a vector. In other words, the vector you are looping
# through is the list where each element of the list is a function
```

### Part B: Applied practice

The mean is defined as

$$
\frac{\sum_{i=1}^n{x_i}}{n}
$$

where $x_i$ represents each element in the mean and $n$ represents the total number of elements. In other words, sum and divide by the number included.

1. Write a function to calculate the mean that removes missing data before
conducting the calculation. You may not use `base::mean` or any similar function that conducts a mean calculation.

```{r ans1, include = FALSE}
my_mean <- function(x) {
	if(any(is.na(x))) {
		tot_na <- sum(is.na(x))
		warning(paste(tot_na, "observations removed due to missingness"))
		x <- na.omit(x)
	}
	sum(x) / length(x)
}
```

2. Test your function to make sure it (a) provides the expected results, and(b)
gives identical output to `base::mean` when `na.rm = TRUE`.

```{r ans2, include = FALSE}
my_mean(mtcars$mpg)
mean(mtcars$mpg)

my_mean(airquality$Ozone)
mean(airquality$Ozone)
mean(airquality$Ozone, na.rm = TRUE)
```

3. Write a function that takes two vectors of the same length and returns
the total number of instances where the value is `NA` for both vectors. For
example, given the following two vectors

```{r vects, eval = FALSE}
c(1, NA, NA, 3, 3, 9, NA)
c(NA, 3, NA, 4, NA, NA, NA)
```

The function should return a value of `2`, because the vectors are both `NA` at
the third and seventh locations. Provide at least one additional test that the function works as expected.

```{r ans3, include = FALSE}
a <- c(1, NA, NA, 3, 3, 9, NA)
b <- c(NA, 3, NA, 4, NA, NA, NA)

sum(is.na(a) & is.na(b))

both_na <- function(x, y) {
	sum(is.na(x) & is.na(y))	
}
both_na(a, b)
both_na(c(rep(NA, 7), rnorm(3)),
        c(rep(NA, 4), rnorm(6)))
```


4. Turn the following into functions. Make sure to give them meaningful names. 
Test them out to make sure they provide the expected output.

```{r practice, eval = FALSE}
mean(is.na(airquality$Ozone))

purrr::map_dbl(split(mtcars$cyl, mtcars$cyl), length) / length(mtcars$cyl)

(mtcars$mpg - mean(mtcars$mpg, na.rm = TRUE)) / 
  sd(mtcars$mpg, na.rm = TRUE) 
```

```{r ans4, include = FALSE}
prop_missing <- function(x) {
	mean(is.na(x))
}
prop_missing(c(rep(NA, 25), rnorm(75)))
prop_missing(c(rep(NA, 42), rnorm(58)))

prop_cat <- function(cat_v) {
	purrr::map_dbl(split(cat_v, cat_v), length) / length(cat_v)
}
prop_cat(rep(letters[1:4], each = 25))
prop_cat(c(rep(letters[1], 50), rep(letters[4:5], each = 25)))


standardize <- function(x) {
	if(any(is.na(x))) {
		tot_na <- sum(is.na(x))
		warning(paste(tot_na, "observations removed due to missingness"))
		x <- na.omit(x)
	}
	(x - mean(x)) / sd(x) 
}
test_standardize <- cbind(standardize(airquality$Ozone), 
                          na.omit(scale(airquality$Ozone)))
identical(test_standardize[ ,1], test_standardize[ ,2])
```

5. Write a function that takes a data frame as its input and returns a data
frame with only the numeric columns.

```{r ans5, include = FALSE}
df_numeric <- function(df, as_tibble = TRUE) {
	num <- purrr::map_lgl(df, is.numeric)
	
	if(as_tibble) {
		return(tibble::as_tibble(df[ ,num]))
	}
	df[ ,num]
}
df_numeric(iris)
df_numeric(iris, as_tibble = FALSE)
```

6. Write a function that uses the function you wrote in Question 5, and returns
a data frame with the mean and standard deviation of each numeric column.

```{r ans6, include = FALSE}
df_summary <- function(df, as_tibble = TRUE) {
	df_num <- df_numeric(df)
	means <- purrr::map_dbl(df_num, my_mean)
	sds <- purrr::map_dbl(df_num, sd, na.rm = TRUE)

	if(as_tibble) {
		tbl <- tibble::tibble(variable = names(df_num),
		               mean = means,
		               sd = sds)
	} else {
		tbl <- data.frame(variable = names(df_num),
		                  mean = means,
		                  sd = sds)
	}
	tbl
}
df_summary(iris)
df_summary(forcats::gss_cat)
```
