---
title: "Lab 4"
author: "Daniel Anderson"
date: "Assigned: 5/8/19; Due 5/15/19"
output:
  html_document: 
    toc: True
    toc_float: true
    theme: "journal"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

There are two parts to this lab. The first includes writing a relatively complicated function. The second involves using that function within a relatively basic shiny app.

## Part A: Writing a summary function
Write a function that takes a given data frame as its first argument, a
grouping variable as its second argument, and a column to summarize as
its third argument. 

### Function behavior
By default, the function should report:

* Total number of observations
* Number of non-missing observations
* Number of missing observations
* Mean
* Standard deviation
* Minimum 
* Maximum 

However, the function should be flexible so it can take **any function(s)** as its input to summarize by, provided the function results in a single value (e.g., anything you could feed to `dplyr::summarize`). You'll need a fourth optional argument that lists the function(s) to use in the summary. 

#### Errors
Embed an error if 

* Anything other than a data frame is supplied as the first
argument
* The provided column to summarize is not numeric. 

#### Warnings
Embed a warning if 

* The grouping variable is numeric (generally you would expect the grouping
variable to be a categorical variable)

### Some guidance
If you want to use {dplyr} you'll need to use non-standard evaluation (NSE), so
something like `!!enquo(group_var)`, where `group_var` is the formal argument
of your function. If you go this route, you will also likely
want to look into `summarize_*`, which take an additional `.funs` argument (to 
make the functions used in summarizing more flexible). Finally, `dplyr::pull` 
(together with NSE) will likely be helpful for checking types for
warnings/errors.

Alternatively, you might consider `split`ing the
dataset by the grouping variable, then looping over each split to apply the
function(s) supplied.

### Sample output
Below are some examples of how the function should work. Note that I have
purposefully given my function a poor name, but you should think carefully about
the name of your function. 

**Examples with {dplyr}:**

```{r include = FALSE}
library(tidyverse)

####### Try playing with the function to get it to work
# group_var <- quo(cyl)
# outcome_var <- quo(mpg)

# mtcars %>%
# 	group_by(!!group_var) %>%
# 	summarize_at(vars(!!outcome_var), list(mean, max))


# Write the actual function
f <- function(data, 
              group_var, 
              outcome_var, 
              .funs = list(n = ~length(.),
                           n_valid = ~sum(!is.na(.)),
                           n_miss = ~sum(is.na(.)),
                           mean = ~mean(., na.rm = TRUE),
                           sd = ~sd(., na.rm = TRUE),
                           min = ~min(., na.rm = TRUE),
                           max = ~max(., na.rm = TRUE))
              ) {
	if(!is.data.frame(data)) {
		stop("This is a non-df error. Make your message more informative")
	}
	if(!is.numeric(pull(data, !!enquo(outcome_var)))) {
		stop("This is an error for the outcome variable being non-numeric. Make your message more informative")
	}
	if(is.numeric(pull(data, !!enquo(group_var)))) {
		warning("This is a warning about a numeric grouping variable. Make your message more informative.")
	}
	data %>%
		group_by(!!enquo(group_var)) %>%
		summarize_at(vars(!!enquo(outcome_var)),
		             .funs)
}
```

```{r examples, error = TRUE}
f(iris, Species, Sepal.Width)
f(airquality, Month, Ozone)
airquality %>%
	mutate(Month = as.factor(Month)) %>%
	f(Month, Ozone)
f(iris$Species, iris$Sepal.Width)

f(iris, Sepal.Width, Species)

f(iris, Species, Sepal.Width, median)
f(iris, Species, Sepal.Width, list(~median(.), 
                                   ~quantile(., 0.2)))
f(iris, Species, Sepal.Width, list(med = ~median(.), 
                                   q20 = ~quantile(., 0.2)))
```

**Examples without {dplyr}**

Note the syntax will look slightly different (because you're not using NSE) but it all should still work basically the same.

```{r f2, include = FALSE}
f2 <- function(data, 
              group_var, 
              outcome_var, 
              .funs = list(n = length,
                           n_valid = function(x) sum(!is.na(x)),
                           n_miss = function(x) sum(is.na(x)),
                           mean = function(x) mean(x, na.rm = TRUE),
                           sd = function(x) sd(x, na.rm = TRUE),
                           min = function(x) min(x, na.rm = TRUE),
                           max = function(x) max(x, na.rm = TRUE))
              ) {
	if(!is.data.frame(data)) {
		stop("This is a non-df error. Make your message more informative")
	}
	if(!is.numeric(data[[outcome_var]])) {
		stop("This is an error for the outcome variable being non-numeric. Make your message more informative")
	}
	if(is.numeric(data[[group_var]])) {
		warning("This is a warning about a numeric grouping variable. Make your message more informative.")
	}
	splt <- split(data, data[[group_var]] )

	map(splt, ~map(.funs, function(f) f(.x[[outcome_var]]))) %>%
		map_df(as.data.frame, .id = "group_var")

}
```

```{r examples-f2}
f2(iris, "Species", "Sepal.Width")
f2(mtcars, "cyl", "mpg")
f2(iris, "Species", "Sepal.Width", list(med = median, 
                                        q20 = function(x) quantile(x, 0.2)))
```

## Part B: Shiny app

Build a shiny app using the `gss_cat` dataset available through the {forcats} package, which you can access with `data(gss_cat, package = "forcats")`. 

The app should allow the user to select a variable, including (potentially) "none", and will produce a histogram of `tvhours` faceted by that variable. Below the plot, render a table summarizing the distribution(s) using the function you wrote in Part A. Note, I would encourage you to take a look at `selectInput` rather than `radioButtons` for the categorical variable, but either should work.

For an example of what the shiny app should look like, see [here](https://djanderson07.shinyapps.io/lab-c3/).